A Higher-Order Approach to Explicit Contexts in LF

### 2015-03-30 paper/paper.tex
typos

  Another technique advanced by Crary is to represent contexts explicitly, and
  then define a sequent judgement over them; this can be used to resolve the
  problem described above (and several others), but it comes at the cost of
  verbosity, and introduces difficulties when used to encode \emph{telescopes}
  in the LF, which are contexts where where each successive term has the
  previous bound variables free in it.  In this paper, I demonstrate a
  higher-order encoding of telescopes which can be used to faithfully encode
  the functional sequent judgement for MLTT~1979.

@@

To motivate this work, let us consider an encoding of a fragment of
Martin-L\"of's extensional, polymorphic type
theory\cite{Martin-Lof-1979} in the logical framework. The syntax
contains two sorts, $\sortval$ (for canonical forms) and $\sortexp$
(for non-canonical forms), and is defined in the following LF
signature:\footnote{LF signatures are presented in the style of
inference rules for clarity. Object judgement forms have a \emph{mode}
which expresses which arguments represent inputs, and which represent
outputs; in our presentation, we color the inputs
\emph{\InputMode\InputModeColorName}, and the outputs
\emph{\OutputMode\OutputModeColorName}. Proper assignment of modes may
be used to construe a judgement as having algorithmic content.}

@@

\section{Encoding Telescopes in Higher-Order Abstract Syntax}
More precisely, it seems that what we really need is an encoding of
\emph{telescopes}, in the sense of De
Bruijn\cite{Bruijn91telescopicmappings}. A telescope is a sequence of
type assignments, where variables bound earlier may appear free later;
for instance:
\[
  \cdot,x_1:A_1, x_2:A_2(x_1), x_3:A_3(x_1,x_2), \dots, x_n:A_n(x_1,\dots,x_{n-1})
\]

@@

Next, we will do something similar in order to show how each of these
judgement(-codes) behaves extensionally, by introducing another algorithmic
judgement $\refract{\dname{J}}{\dname{J}^=}$. Intuitively, this judgement
describes how what happens when you substitute two purportedly equal
expressions in for the judgement-code's outermost variable.

@@

The judgement $\sploot{\dname{J}}{\dname{J}'}$ may be considered as
defining the behavior of $\dname{J}$ with one hole punched in it;
$[x]\;\dname{J}(x)$ could be called the ``one-hole application of
$\dname{J}$''. Likewise, $\refract{\dname{J}}{\dname{J}^=}$ may be
thought of as defining the behavior $\dname{J}$ with two hole punched
in it, and $[x][y]\; \dname{J}^=(x,y)$ could be called the ``two-hole
application of $\dname{J}$''.

>>>>

  Another technique advanced by Crary is to represent contexts explicitly, and
  then define a sequent judgement over them; this can be used to resolve the
  problem described above (and several others), but it comes at the cost of
  verbosity, and introduces difficulties when used to encode \emph{telescopes}
  in the LF, which are contexts where each successive term has the
  previous bound variables free in it.  In this paper, I demonstrate a
  higher-order encoding of telescopes which can be used to faithfully encode
  the functional sequent judgement for MLTT~1979.

@@

To motivate this work, let us consider an encoding of a fragment of
Martin-L\"of's extensional, polymorphic type
theory \cite{Martin-Lof-1979} in the logical framework. The syntax
contains two sorts, $\sortval$ (for canonical forms) and $\sortexp$
(for non-canonical forms), and is defined in the following LF
signature:\footnote{LF signatures are presented in the style of
inference rules for clarity. Object judgement forms have a \emph{mode}
which expresses which arguments represent inputs, and which represent
outputs; in our presentation, we color the inputs
\emph{\InputMode\InputModeColorName}, and the outputs
\emph{\OutputMode\OutputModeColorName}. Proper assignment of modes may
be used to construe a judgement as having algorithmic content.}

@@

\section{Encoding Telescopes in Higher-Order Abstract Syntax}
More precisely, it seems that what we really need is an encoding of
\emph{telescopes}, in the sense of De
Bruijn \cite{Bruijn91telescopicmappings}. A telescope is a sequence of
type assignments, where variables bound earlier may appear free later;
for instance:
\[
  \cdot,x_1:A_1, x_2:A_2(x_1), x_3:A_3(x_1,x_2), \dots, x_n:A_n(x_1,\dots,x_{n-1})
\]

@@

Next, we will do something similar in order to show how each of these
judgement(-codes) behaves extensionally, by introducing another algorithmic
judgement $\refract{\dname{J}}{\dname{J}^=}$. Intuitively, this judgement
describes what happens when you substitute two purportedly equal
expressions in for the judgement-code's outermost variable.

@@

The judgement $\sploot{\dname{J}}{\dname{J}'}$ may be considered as defining
the behavior of $\dname{J}$ with one hole punched in it; $[x]\;\dname{J}(x)$
could be called the ``one-hole application of $\dname{J}$''. Likewise,
$\refract{\dname{J}}{\dname{J}^=}$ may be thought of as defining the behavior
$\dname{J}$ with two holes punched in it, and $[x][y]\; \dname{J}^=(x,y)$ could
be called the ``two-hole application of $\dname{J}$''.
